---
title: "Fig5"
author: "Saye Kazem"
output: github_document
---

## Fig. 5: Rare and common variant architectures across complex traits.

#### -- Figure legend -- ####
**Legend:** (A) Example of genetic and phenotypic correlations between fluid intelligence (FI) and 6 traits, with each point color-coded by correlation type: SNP-based genetic correlations, deletion, duplication burden correlations, loss-of-function SNV burden correlation, and phenotypic correlations. (B) The heatmap of burden correlations, for deletions and duplications in the lower and upper triangle matrices, respectively. Significant correlations are marked by asterisks (FDR corrected p-Jaccard). (C) Network graph illustrating the mean absolute correlations for deletions and duplications across five distinct functional trait categories. Node size reflects the number of traits in each category, and edge thickness corresponds to the correlation strength. (D) Scatter plot comparing effect size correlations between deletions and duplications, where orange lines represent the Total Least Squares regression. The dashed line indicates perfect concordance (y = x). (E) Boxplots illustrate the variance explained (heritability, R2) by gene sets for both brain and non-brain traits, based on regression analysis. (F-G) Scatter plots comparing CNV burden correlations (deletions and duplications) with SNP-based genetic correlations. (H-I) Scatter plots comparing CNV burden correlations with SNV burden correlations. Abbreviations: BMI: body mass index; EA: educational attainment; FI: fluid intelligence; TDI: townsend deprivation index; VSWM: visuospatial working memory; FVC: Forced vital capacity; Del: deletion; Dup: duplication; SNP: single nucleotide polymorphism; SNV: single nucleotide variants; LoF: loss of function.

#### Libraries ####
```{r setup, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggcorrplot)
library(ggprism)
library(ggpubr)
library(MASS)
library(dplyr)
library(tidyr)
library(reshape2)
library(Rmpfr)
library(vegan)
library(openxlsx)
library(svglite)
library(igraph)
library(ggraph)
library(forcats)
library(knitr)
library(patchwork)

# Define Core Colors
brain_colorcode  <- "orchid4"      
nonbrain_colorcode <- "gold3"  
in_base_size_ggprism <- 14

message("Loading Figure 5 Data...")
load("df_betweentrait_correlations.RData")
load("df_traitlevel_stats.RData")
load("df_tls_regression.RData") 

# Helper Function: Total Least Squares
total_least_squares <- function(X, Y) {
  X <- as.matrix(X)
  Y <- as.matrix(Y)
  Z <- cbind(X, Y)
  svd_result <- svd(Z)
  V <- svd_result$v
  nx <- ncol(X)
  Vxy <- V[1:nx, (nx + 1):ncol(V)]
  Vyy <- V[(nx + 1):ncol(V), (nx + 1):ncol(V)]
  return(-Vxy %*% solve(Vyy))
}

# Robust function to calculate TLS slope and its p-value
compute_tls_stats <- function(df, x_var, y_var) {
  X <- df[[x_var]]; Y <- df[[y_var]]
  
  # 1. Slope
  slope <- as.numeric(total_least_squares(X, Y))
  
  # 2. P-value (Testing deviation from slope = 0)
  n <- length(X)
  mean_x <- mean(X); mean_y <- mean(Y)
  s_xx <- sum((X - mean_x)^2); s_yy <- sum((Y - mean_y)^2)
  s_xy <- sum((X - mean_x) * (Y - mean_y))
  
  std_err <- sqrt((s_yy - slope * s_xy) / (n - 2)) / sqrt(s_xx)
  t_stat <- slope / std_err
  
  # Calculate p-value
  p_val <- 2 * (1 - pt(abs(t_stat), df = n - 2))
  
  return(data.frame(Slope = slope, P_Value = p_val))
}

build_scatter <- function(df, x_var, y_var, x_lab, y_lab) {
  slope <- total_least_squares(df[[x_var]], df[[y_var]])
  
  ggplot(df, aes_string(x = x_var, y = y_var)) + 
    geom_point(color = "grey", alpha=0.4) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "darkgrey") +
    geom_abline(slope = slope, intercept = 0, color = "orange", linewidth=1) +
    coord_cartesian(ylim = c(-1, 1), xlim = c(-1, 1)) +
    labs(x = x_lab, y = y_lab) +
    theme_prism(border = FALSE, base_size = 14, base_fontface = "bold") + 
    theme(axis.line = element_line(linewidth = 0.8))
}
```

#### Panel A: Correlation with FI (Lollipop Chart)
Description: Lollipop chart illustrating the correlation between fluid intelligence (FI) and six selected traits across multiple genetic and phenotypic correlation metrics. ####

##### Panel A Data #####
```{r, fig.width=7, fig.height=3 ,dpi=100 }
load("df_betweentrait_correlations.Rdata")
FI_Others_plot <- df_betweentrait_correlations %>%
  filter(Trait1 == "FI" | Trait2 == "FI") %>%
  mutate(Trait2 = ifelse(Trait1 == "FI", Trait2, Trait1)) %>%
  dplyr::select(Trait2, Corr_DEL, Corr_DUP, Corr_SNV, Corr_SNP, Corr_Pheno) %>% # Fix applied here
  pivot_longer(cols = starts_with("Corr_"), names_to = "correlation_type", values_to = "correlation") %>%
  mutate(correlation_type = case_when(
    correlation_type == "Corr_DEL" ~ "DEL-burden",
    correlation_type == "Corr_DUP" ~ "DUP-burden",
    correlation_type == "Corr_SNV" ~ "SNV-lof-burden",
    correlation_type == "Corr_SNP" ~ "SNP (rg)",
    correlation_type == "Corr_Pheno" ~ "Phenotype"
  )) %>%
  filter(Trait2 %in% c("Reaction", "Neuroticism", "TDI", "Redblood", "BMI", "StandingHeight")) %>%
  mutate(Trait2 = fct_recode(factor(Trait2), "Height" = "StandingHeight")) %>%
  mutate(Trait2 = factor(Trait2, levels = c("Reaction", "Neuroticism", "TDI", "Redblood", "BMI", "Height"))) %>%
  mutate(correlation_type = factor(correlation_type, levels = c("Phenotype","DUP-burden","SNP (rg)","DEL-burden","SNV-lof-burden")))
```

##### Panel A Figure #####
```{r, fig.width=5, fig.height=8, dpi=250}
pA_lollipop <- ggplot(FI_Others_plot, aes(x=correlation, y=Trait2, color=correlation_type)) +
  geom_errorbar(aes(xmin = 0, xmax = correlation), position = position_dodge(width = 0.5), height = 0, alpha=0.7) +
  geom_point(size=2, position = position_dodge(width = 0.5), alpha=0.8) +
  geom_vline(lty = 2, xintercept = 0, color='darkgrey') +
  scale_color_manual(values=c("DEL-burden" = "red3", "DUP-burden" = "royalblue", "SNV-lof-burden" = "orange3", "SNP (rg)" = "green4", "Phenotype" = "purple4")) +
  labs(x = "Correlation with FI", y = "", color = "Correlation Type") +
  theme_prism(base_size = in_base_size_ggprism) + 
  theme(panel.grid.major.y = element_blank(), panel.border = element_blank(), axis.ticks.y = element_blank(), axis.line = element_line(linewidth = 0.8))

print(pA_lollipop)
```


##### Panel A Stats #####
```{r, fig.width=18, fig.height=7 ,dpi=800 }
kable(head(FI_Others_plot), format = "markdown", caption = "Extracted Correlations for Fluid Intelligence")
```


#### Panel B: Burden Correlation Heatmap
Description: Heatmap displaying pairwise burden correlations for deletions (lower triangle) and duplications (upper triangle) across a subset of 14 complex traits.

##### Panel B Data #####
```{r, fig.width=18, fig.height=7 ,dpi=800 }
# 1. Trait Subset
SUBSET_Order <- c("HDL", "FVC", "EA", "FI", "Menopause", "Menarche", 
                  "Depression", "Neuroticism", "Loneliness", "MoodAnxiety", "GuiltyFeelings", 
                  "TDI", "VSWM1", "BMI", "HeartRate", "Hypertention", 
                  "Triglycerides", "Platelet", "Redblood", "Gamma")

# Filter data to the subset
df_heatmap <- df_betweentrait_correlations %>%
  filter(Trait1 %in% SUBSET_Order & Trait2 %in% SUBSET_Order)

# 2. Initialize empty symmetric matrices
Corr_Del_matrix <- matrix(NA, nrow = length(SUBSET_Order), ncol = length(SUBSET_Order), dimnames = list(SUBSET_Order, SUBSET_Order))
Corr_Dup_matrix <- matrix(NA, nrow = length(SUBSET_Order), ncol = length(SUBSET_Order), dimnames = list(SUBSET_Order, SUBSET_Order))
CorrPval_Del_matrix <- matrix(NA, nrow = length(SUBSET_Order), ncol = length(SUBSET_Order), dimnames = list(SUBSET_Order, SUBSET_Order))
CorrPval_Dup_matrix <- matrix(NA, nrow = length(SUBSET_Order), ncol = length(SUBSET_Order), dimnames = list(SUBSET_Order, SUBSET_Order))

# 3. Fill in the matrices explicitly 
for(i in 1:nrow(df_heatmap)){
  t1 <- as.character(df_heatmap$Trait1[i])
  t2 <- as.character(df_heatmap$Trait2[i])
  
  # Correlations
  Corr_Del_matrix[t1, t2] <- df_heatmap$Corr_DEL[i]
  Corr_Del_matrix[t2, t1] <- df_heatmap$Corr_DEL[i]
  
  Corr_Dup_matrix[t1, t2] <- df_heatmap$Corr_DUP[i]
  Corr_Dup_matrix[t2, t1] <- df_heatmap$Corr_DUP[i]
  
  # P-values (FDR Jaccard)
  CorrPval_Del_matrix[t1, t2] <- df_heatmap$Corr_FDR_JaccardPval_DEL[i]
  CorrPval_Del_matrix[t2, t1] <- df_heatmap$Corr_FDR_JaccardPval_DEL[i]
  
  CorrPval_Dup_matrix[t1, t2] <- df_heatmap$Corr_FDR_JaccardPval_DUP[i]
  CorrPval_Dup_matrix[t2, t1] <- df_heatmap$Corr_FDR_JaccardPval_DUP[i]
}

# 4. Enforce the exact specified ordering on all matrices
Corr_Del_matrix <- Corr_Del_matrix[SUBSET_Order, SUBSET_Order]
Corr_Dup_matrix <- Corr_Dup_matrix[SUBSET_Order, SUBSET_Order]
CorrPval_Del_matrix <- CorrPval_Del_matrix[SUBSET_Order, SUBSET_Order]
CorrPval_Dup_matrix <- CorrPval_Dup_matrix[SUBSET_Order, SUBSET_Order]

# 5. Combine Correlations: Lower = DEL, Upper = DUP
combined_corr_ordered_subset <- Corr_Del_matrix
combined_corr_ordered_subset[upper.tri(combined_corr_ordered_subset)] <- Corr_Dup_matrix[upper.tri(Corr_Dup_matrix)]
diag(combined_corr_ordered_subset) <- Inf

# 6. Combine P-values: Lower = DEL, Upper = DUP (Ensures stars match the colors!)
combined_pval_ordered_subset <- CorrPval_Del_matrix
combined_pval_ordered_subset[upper.tri(combined_pval_ordered_subset)] <- CorrPval_Dup_matrix[upper.tri(CorrPval_Dup_matrix)]
diag(combined_pval_ordered_subset) <- NA # Prevent stars on the diagonal

# 7. Convert P-values to stars
significance_levels <- function(p) {
  ifelse(is.na(p), "", ifelse(p < 0.05, "*", ""))
}
stars_matrix_subset <- matrix(significance_levels(combined_pval_ordered_subset), 
                              nrow = nrow(combined_pval_ordered_subset),
                              dimnames = list(SUBSET_Order, SUBSET_Order))
kable(head(combined_corr_ordered_subset), format = "markdown", caption = "Corr Matrix")
```


##### Panel B Figure #####
```{r, fig.width=7, fig.height=7, dpi=400}
gap_indices <- c(1, 2, 4, 6, 11, 13, 16)
line_positions <- gap_indices + 0.5

# Base correlation plot
pB_heatmap <- ggcorrplot(combined_corr_ordered_subset, 
                         type = "full", 
                         colors = c("darkblue", "white", "darkred"),
                         show.diag = TRUE)

# Add white separator lines for grouped blocks
pB_heatmap <- pB_heatmap + 
  geom_vline(xintercept = line_positions, color = "white", linewidth = 1, linetype = "solid") +
  geom_hline(yintercept = line_positions, color = "white", linewidth = 1, linetype = "solid") 
  #theme_prism(axis_text_angle = 45, border = FALSE, base_size = 14, base_fontface = "bold")

# Overlay the correctly aligned FDR significance stars
for (i in 1:nrow(combined_corr_ordered_subset)) {
  for (j in 1:ncol(combined_corr_ordered_subset)) {
    if (!is.na(combined_corr_ordered_subset[i, j]) && stars_matrix_subset[i, j] == "*") {
      pB_heatmap <- pB_heatmap + annotate("text", x = i, y = j - 0.2, label = stars_matrix_subset[i, j], color = "grey38", size = 9, fontface = "bold")
    }
  }
}

print(pB_heatmap)
```

##### Panel B Stats #####
```{r, fig.width=9, fig.height=9, dpi=400}
# Display summary of FDR p-values used for heatmap annotations
# Display summary of FDR p-values used for heatmap annotations
kable(head(df_heatmap %>% dplyr::select(Trait1, Trait2, Corr_FDR_JaccardPval_DEL, Corr_FDR_JaccardPval_DUP)), format = "markdown", caption = "FDR Jaccard P-values for Categorized Heatmap")
```
#### Panel C: Mean Absolute Correlation Network ####
Description: Spring-embedded network graph showing the mean absolute burden correlations among five major trait categories for deletions and duplications. Node size/alpha reflects the number of traits; edge width/alpha reflects correlation strength.

##### Panel C Data & Stats #####
```{r, fig.width=9, fig.height=9, dpi=400}
# 1. Define trait categories (nodes)
groups_list <- list(
  'Blood Assays' = c('Albumin', 'IGF1', 'Cholesterol', 'HDL', 'LDL', 'Calcium', 'Platelet', 'Redblood', 'Gamma', 'HbA1c', 'Triglycerides', 'Glucose'),
  'Reproductive and Activity Factors' = c('Menarche', 'Menopause', 'SexPartners', 'MET'),
  'Physical Measure' = c('FVC', 'FEV1', 'StandingHeight', 'BirthWeight', 'BMI', 'BMD', 'Hypertention', 'HeartRate'),
  'Cognitive Metrics' = c('EA', 'FI', 'SymbolDigit', 'TrailMaking1', 'TrailMaking2', 'Reaction', 'VSWM1', 'VSWM2', 'TDI'),
  'Mental Health' = c('RiskTaking', 'Depression', 'GP', 'Neuroticism', 'Irritability', 'MoodSwings', 'Miserableness', 'GuiltyFeelings', 'Loneliness', 'MoodAnxiety')
)

trait_to_group <- stack(groups_list) %>% setNames(c("Trait", "Group")) %>% tibble::deframe()

# 2. Filter and map traits to groups using the unified correlation dataframe
Net_Corr <- df_betweentrait_correlations %>%
  filter(!is.na(Corr_DEL), !is.na(Corr_DUP), !is.na(Trait1), !is.na(Trait2)) %>%
  mutate(
    Group1 = trait_to_group[Trait1],
    Group2 = trait_to_group[Trait2]
  )

node_sizes <- data.frame(name = names(groups_list), Size = sapply(groups_list, length))

# 3. Combined Edge Calculation for Mean Absolute Correlation
calculate_mean_abs_corr <- function(g1, g2, corr_data, cnv_type) {
  corr_col <- paste0("Corr_", cnv_type)
  subset_df <- corr_data %>%
    filter((Group1 == g1 & Group2 == g2) | (Group1 == g2 & Group1 == g1))
  
  abs_corrs <- abs(subset_df[[corr_col]])
  mean_abs_corr <- mean(abs_corrs, na.rm = TRUE)
  
  if (is.nan(mean_abs_corr) || is.null(mean_abs_corr)) return(NULL)
  return(data.frame(from = g1, to = g2, Metric = mean_abs_corr))
}

group_pairs <- t(combn(names(groups_list), 2))
del_edges_corr <- do.call(rbind, apply(group_pairs, 1, function(row) calculate_mean_abs_corr(row[1], row[2], Net_Corr, "DEL")))
dup_edges_corr <- do.call(rbind, apply(group_pairs, 1, function(row) calculate_mean_abs_corr(row[1], row[2], Net_Corr, "DUP")))

del_edges_corr$CNV_Type <- "DEL"
dup_edges_corr$CNV_Type <- "DUP"

all_edges <- rbind(del_edges_corr, dup_edges_corr)

# Extract scaling limits for the plot
global_max_metric <- max(all_edges$Metric, na.rm = TRUE)
min_node_size_val <- min(node_sizes$Size, na.rm = TRUE)
max_node_size_val <- max(node_sizes$Size, na.rm = TRUE)
kable(head(all_edges), format = "markdown", caption = "Corr Matrix")

```

##### Panel C Figure #####
```{r, fig.width=20, fig.height=16, dpi=200}
# 1. Create a single graph object with all edges
graph_combined <- graph_from_data_frame(d = all_edges, vertices = node_sizes, directed = FALSE)

# 2. Create the weighted layout
set.seed(42)
layout <- create_layout(graph_combined, layout = 'fr', weights = E(graph_combined)$Metric)

# 3. Build the publication-ready plot
pC_network <- ggraph(layout) +
  geom_edge_fan(
    aes(
      color = CNV_Type,
      width = Metric,
      label = round(Metric, 2),
      alpha = Metric # Alpha mapped to Metric for edges
    ),
    label_dodge = unit(4, 'mm'),
    angle_calc = 'along',
    label_colour = 'grey20',
    label_size = 8
  ) +
  geom_node_point(
    aes(size = Size, alpha = Size), # Alpha mapped to Size
    color = "grey20" 
  ) +
  geom_node_label(
    aes(label = name), 
    repel = TRUE, 
    size = 8, 
    fontface = "bold", 
    fill = alpha("white", 0.9), 
    box.padding = unit(2, "lines"), 
    label.r = unit(0.2, "lines"), 
    label.size = 0.6 
  ) +
  # Scales
  scale_edge_width(
    range = c(1, 8),
    name = "Mean Absolute\nCorrelation ($|\\bar{r}|$)",
    limits = c(0, global_max_metric)
  ) +
  scale_edge_color_manual(
    name = "CNV Type",
    values = c("DEL" = "darkred", "DUP" = "darkblue")
  ) +
  scale_edge_alpha_continuous(
    name = "Mean Absolute\nCorrelation ($|\\bar{r}|$)",
    range = c(0.2, 0.9),
    limits = c(0, global_max_metric),
    guide = "none"
  ) +
  scale_size_continuous(
    range = c(10, 30),
    name = "Number of Traits"
  ) +
  scale_alpha_continuous(
    name = "Number of Traits", 
    range = c(0.4, 0.9),       
    limits = c(min_node_size_val, max_node_size_val),
    guide = "legend"           
  ) +
  scale_fill_manual(
    values = c("white"),
    guide = "none" 
  ) +
  theme_graph(base_family = 'sans') +
  labs(title = "Combined Network of DEL and DUP Trait Associations\n(Mean Absolute Correlation)") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
    legend.position = 'right',
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 16)
  )

print(pC_network)
```


##### Panel D Data #####

```{r, fig.width=5, fig.height=5, dpi=200}
df_pD <- df_betweentrait_correlations %>% 
  filter(Include_In_DEL_DUP_Corr == TRUE) %>% 
  drop_na(Corr_DUP, Corr_DEL)

# Calculate the Total Least Squares (TLS) slope
slope_pD <- total_least_squares(df_pD$Corr_DUP, df_pD$Corr_DEL)
kable(head(df_pD), format = "markdown", caption = "Del-Dup Burden Corr")
```

##### Panel D Figure #####

```{r, fig.width=4, fig.height=4, dpi=180}
pD_scatter <- ggplot(df_pD, aes(x = Corr_DUP, y = Corr_DEL)) + 
  geom_point(color = "grey", alpha = 0.4) +
  # Dashed line represents perfect concordance (y = x)
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "darkgrey") +
  # Orange line represents the TLS regression fit
  geom_abline(slope = slope_pD, intercept = 0, color = "orange", linewidth = 1) +
  coord_cartesian(ylim = c(-1, 1), xlim = c(-1, 1)) +
  labs(x = "Burden Correlation - Dup", y = "Burden Correlation - Del") +
  theme_prism(border = FALSE, base_size = 14, base_fontface = "bold") + 
  theme(axis.line = element_line(linewidth = 0.8))

print(pD_scatter)
```

##### Panel D Stats #####

```{r, fig.width=4, fig.height=4, dpi=180}
stats_pD <- compute_tls_stats(df_pD, "Corr_DUP", "Corr_DEL")

# 2. Display in a clean table
kable(stats_pD, format = "markdown", caption = "TLS Regression Statistics (Dup vs Del)")
```



##### Panel E Data #####
```{r, fig.width=4, fig.height=4, dpi=180}
# Reshape the trait-level stats into long format for plotting
df_variance <- df_traitlevel_stats %>%
  dplyr::select(Trait, Trait_Cat, Del = Del_VarianceExplained, Dup = Dup_VarianceExplained) %>%
  pivot_longer(cols = c(Del, Dup), names_to = "CNV_Type", values_to = "Heritability") %>%
  drop_na(Heritability) %>%
  # Ensure clean naming for the legend and correct factor order
  mutate(Trait_Cat = ifelse(Trait_Cat == "Brain", "Brain", "Non-Brain")) %>%
  mutate(Trait_Cat = factor(Trait_Cat, levels = c("Brain", "Non-Brain")),
         CNV_Type = factor(CNV_Type, levels = c("Del", "Dup")))
kable(head(subset(df_variance,select=c('Trait', 'Trait_Cat','CNV_Type','Heritability'))), format = "markdown", caption = "Variance Explained")
```

##### Panel E Figure #####
```{r, fig.width=4, fig.height=4, dpi=180}
pE_variance <- ggplot(df_variance, aes(x = CNV_Type, y = Heritability, fill = Trait_Cat)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.82) +
  geom_jitter(
    size = 0.8,
    alpha = 1,
    shape = 21,
    stroke = 0.4,
    position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.5)
  ) + 
  scale_fill_manual(values = c("Brain" = "orchid4", "Non-Brain" = "gold3")) +
  labs(x = "", y = "Variance explained, R2\n (%)") +
  theme_prism(border = FALSE, base_size = 14, base_fontface = "bold") +
  theme(
    legend.position = "none",
    axis.line = element_line(linewidth = 0.8)
  ) + 
  ylim(0, 0.2)

print(pE_variance)
```

##### Panel E Stats #####
```{r, fig.width=4, fig.height=4, dpi=180}
# 1. Compute Wilcoxon Rank Sum tests for Del and Dup separately
wilcox_del <- wilcox.test(Heritability ~ Trait_Cat, data = subset(df_variance, CNV_Type == "Del"))
wilcox_dup <- wilcox.test(Heritability ~ Trait_Cat, data = subset(df_variance, CNV_Type == "Dup"))

# 2. Adjust p-values using FDR
pvals <- c(wilcox_del$p.value, wilcox_dup$p.value)
pvals_fdr <- p.adjust(pvals, method = "fdr")

# 3. Compile into a clean statistics table
stats_panel_e <- data.frame(
  CNV_Type = c("Del", "Dup"),
  Test = "Wilcoxon Rank Sum Test",
  Comparison = "Brain vs Non-Brain",
  W_Statistic = c(wilcox_del$statistic, wilcox_dup$statistic),
  P_Value = pvals,
  FDR_Q_Value = pvals_fdr
)

kable(stats_panel_e, format = "markdown", caption = "Heritability Variance Explained Stats")
```



##### Panel F & G  Data #####
```{r, fig.width=4, fig.height=4, dpi=180}
# F & G: SNP vs DEL/DUP (Intersection of valid pairs)
df_pFG <- df_betweentrait_correlations %>% 
  filter(Include_In_SNP_DEL_Corr == TRUE & Include_In_SNP_DUP_Corr == TRUE) %>% 
  drop_na(Corr_SNP, Corr_DEL, Corr_DUP)
kable(head(df_pFG), format = "markdown")

```

##### Panel F & G Figure #####
```{r, fig.width=8, fig.height=4, dpi=180}
pF_scatter <- build_scatter(df_pFG, "Corr_SNP", "Corr_DEL", "Genetic Correlation - SNP", "Burden Correlation - Del")
pG_scatter <- build_scatter(df_pFG, "Corr_SNP", "Corr_DUP", "Genetic Correlation - SNP", "Burden Correlation - Dup")

print(pF_scatter | pG_scatter)

```

##### Panel F & G Stats #####
```{r, fig.width=4, fig.height=4, dpi=180}
stats_pF <- compute_tls_stats(df_pFG, "Corr_SNP", "Corr_DEL")
stats_pG <- compute_tls_stats(df_pFG, "Corr_SNP", "Corr_DUP")

stats_fg <- bind_rows(stats_pF, stats_pG) %>% 
  mutate(Panel = c("Panel F (SNP vs DEL)", "Panel G (SNP vs DUP)")) %>%
  dplyr::select(Panel, Slope, P_Value) 

kable(stats_fg, format = "markdown", caption = "TLS Regression Statistics (SNP vs CNV)")
```


##### Panel H & I  Data #####

```{r, fig.width=4, fig.height=4, dpi=180}
# H & I: SNV vs DEL/DUP (Intersection of valid pairs)
df_pHI <- df_betweentrait_correlations %>% 
  filter(Include_In_SNV_DEL_Corr == TRUE & Include_In_SNV_DUP_Corr == TRUE) %>% 
  drop_na(Corr_SNV, Corr_DEL, Corr_DUP)
kable(head(df_pHI), format = "markdown")
```

##### Panel H & I  Figure #####

```{r, fig.width=8, fig.height=4, dpi=180}
library(patchwork)

pH_scatter <- build_scatter(df_pHI, "Corr_SNV", "Corr_DEL", "Burden Correlation - SNV", "Burden Correlation - Del")
pI_scatter <- build_scatter(df_pHI, "Corr_SNV", "Corr_DUP", "Burden Correlation - SNV", "Burden Correlation - Dup")

print(pH_scatter | pI_scatter)
```

##### Panel H & I  Stats #####
```{r, fig.width=8, fig.height=4, dpi=180}
stats_pH <- compute_tls_stats(df_pHI, "Corr_SNV", "Corr_DEL")
stats_pI <- compute_tls_stats(df_pHI, "Corr_SNV", "Corr_DUP")

stats_hi <- bind_rows(stats_pH, stats_pI) %>% 
  mutate(Panel = c("Panel H (SNV vs DEL)", "Panel I (SNV vs DUP)")) %>%
  dplyr::select(Panel, Slope, P_Value)

kable(stats_hi, format = "markdown", caption = "TLS Regression Statistics (SNV vs CNV)")

```
