---
title: "Fig3"
author: "Sayeh Kazem"
output: github_document
---

## Fig. 3: Whole-body cell type-specific associations of CNVs with complex traits.

#### -- Figure legend -- ####
**Legend:** (A) Heatmap displays a representative set of association effect sizes between five categories of traits (x-axis) and deletions and duplications aggregated by whole-body cell types (y-axis). Traits are categorized on the x-axis, and gene sets are listed on the y-axis with annotations on the right. Blue and red intensity reflects negative and positive effect sizes. Asterisks (*) indicate FDR-significant associations (two-sided t-test from multiple linear regression; 172 gene sets X 43 traits X 2 CNV types = 14,792 tests). A representative subset is shown (complete heatmaps in Figs. S28-S29). 
(B-C) Summary of variant association differences across cell types. (B) Whole-Body Comparison: Differences in aggregate association/enrichment levels across variant types for all cell type gene sets. (C) Higher pleiotropy in brain cell types: Proportion of significant trait associations for brain and non-brain cell type gene sets. In both panels, data are presented as the aggregate statistical proportion of significant enrichments ± standard error. Statistics are derived from gene sets rather than individual sample observations. For these panels, n = 81 whole-body cell type gene sets (n = 7 brain, n = 74 non-brain). No technical or biological replicates are utilized. Asterisks (*) denote FDR-adjusted statistical significance (two-sided Pearson’s Chi-squared test). 
(D) Pearson correlation of gene-set effect sizes between AoU and UKBB. The two-sided Pearson correlation test demonstrates replication for both deletions (r = 0.70, p = 3.74e-13) and duplications (r = 0.41, p = 1.20e-04). The solid line denotes the linear regression fit, with shaded regions representing the 95% confidence intervals, alongside a dashed line representing perfect concordance (y = x). 
(E) Discovery power and cross-cohort replication. Per-trait concordance of 81 gene-set effect sizes (BMI, Heart Rate, Platelet Count, Standing Height) vs. number of UKBB FDR-significant gene sets (Pearson correlation test, r = 0.75, p = 3.30e-02). 
Abbreviations: HDL, high-density lipoprotein; HbA1c, glycated haemoglobin; BMD, bone mineral density; BMI, body mass index; EA, educational attainment; FI, fluid intelligence; TDI, Townsend deprivation index; VSWM, visuospatial working memory; FVC, forced vital capacity; Del, deletion; Dup, duplication; SNP, single nucleotide polymorphism; AoU, All of Us.

#### Libraries ####
```{r setup, message=FALSE, warning=FALSE}
#### Libraries for Heatmap, Bar-Plots, and Exporting
library(ggplot2)
library(ggprism)
library(grid)
library(ggpubr)
library(gridExtra)
library(tidyr)
library(tibble)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
library(svglite) 
library(knitr)   
library(patchwork)
library(stringr)
library(ggrepel)
```


#### Panel A: Cell type-specific associations of CNVs with complex traits (Heatmap)  ####
##### Description: The heatmap displays association effect sizes between five categories of traits (x-axis) and cell type-specific gene sets (y-axis) for deletions (left) and duplications (right). ####


##### Panel A Data #####

```{r}
# Load the FunBurd associations output for UKBB dataset 
load('df_funburd_ukbb.RData')

# Render a clean markdown table preview for GitHub
kable(head(df_funburd_ukbb), format = "markdown")
```


##### Panel A Figure (For Deletion) ######

```{r ,  fig.width=14, fig.height=15}
# ==============================================================================
# 1. DEFINE VECTORS FOR CUSTOM ORDERING (32 Representative Cell Types x 20 Traits)
# ==============================================================================
cell_type_order <- c(
  # Brain Tissues (7)
  "inhibitory_neurons", "excitatory_neurons",
  "astrocytes", "oligodendrocytes", "oligodendrocyte_precursor", "microglial", "muller_glia",
  # Non-Brain Tissues (25)
  "t_cells", "b_cells", "monocytes", "macrophages",
  "endothelial", "lymphatic_endothelial",
  "skeletal_myocytes", "cardiomyocytes",
  "leydig", "enteroendocrine",
  "mesothelial", "hepatocytes",
  "exocrine_glandular", "paneth", "basal_prostatic", "breast_myoepithelial", "ciliated",
  "spermatogonia", "early_spermatids",
  "squamous_epithelial", "suprabasal_keratinocytes",
  "syncytiotrophoblasts", "extravillous_trophoblasts",
  "endometrial_stromal", "peritubular"
)

trait_order <- c(
  # Blood Assays (5)
  'HDL', 'Platelet', 'Redblood', 'Gamma', 'Triglycerides',
  # Reproductive Factors (2)
  'Menarche', 'Menopause',
  # Physical Measures (4)
  'FVC', 'BMI', 'Hypertention', 'HeartRate',
  # Cognitive Metrics (4)
  'EA', 'FI', 'VSWM1', 'TDI',
  # Mental Health (5)
  'Depression', 'Neuroticism', 'GuiltyFeelings', 'Loneliness', 'MoodAnxiety'
)

# ==============================================================================
# 2. SUBSETTING THE MASTER DATAFRAME
# ==============================================================================
# IMPORTANT: Adjust 'CellType_HPA81' if your dataframe uses a different exact string
df_HPACellType = subset(df_funburd_ukbb, Geneset_Type == 'SC_HPA81')
df_HPACellType_Del = subset(df_HPACellType, CNV_Type == 'DEL')

df_HPACellType_Del_EffectSize = subset(df_HPACellType_Del, select = c('Trait', 'Geneset', 'Effectsize'))
df_HPACellType_Del_fdrpval = subset(df_HPACellType_Del, select = c('Trait', 'Geneset', 'FDR_p_value'))

# ==============================================================================
# 3. DATA TRANSFORMATION (LONG TO WIDE FORMAT)
# ==============================================================================
Matrix_HPACellType_Del_EffectSize <- df_HPACellType_Del_EffectSize %>%
  pivot_wider(names_from = Trait, values_from = Effectsize) %>%
  column_to_rownames("Geneset")

Matrix_HPACellType_Del_fdrpval <- df_HPACellType_Del_fdrpval %>%
  pivot_wider(names_from = Trait, values_from = FDR_p_value) %>%
  column_to_rownames("Geneset")

# ==============================================================================
# 4. ORDERING AND SAFE SUBSETTING
# ==============================================================================
Matrix_HPACellType_Del_EffectSize_ordered <- Matrix_HPACellType_Del_EffectSize[cell_type_order, trait_order]
Matrix_HPACellType_Del_fdrpval_ordered <- Matrix_HPACellType_Del_fdrpval[cell_type_order, trait_order]

Matrix_HPACellType_Del_fdrpval_ordered <- ifelse(
  !is.na(Matrix_HPACellType_Del_fdrpval_ordered) & Matrix_HPACellType_Del_fdrpval_ordered < 0.05, 
  "*", 
  ""
)

# ==============================================================================
# 5. HEATMAP COLOR PALETTE SETUP
# ==============================================================================
colors <- colorRampPalette(c("#386cb0", "white", "#a50f15"))(100)
min_val <- -0.2
breaks <- seq(min_val, -min_val, length.out = length(colors))
color_function <- colorRamp2(breaks, colors)

# ==============================================================================
# 6. ANNOTATIONS AND GROUPING
# ==============================================================================
## ROW ANNOTATIONS (Must be exactly 32 elements to match the forced matrix rows)
row_annotation <- data.frame(
  Tissue = c(rep("Brain Cell Types", 7), rep("Non-Brain Cell Types", 25)),
  Cat_Detailed = c(
    rep("Neuronal", 2), rep("Glial", 5), 
    rep("Blood & Immune", 4), rep("Endothelial", 2), 
    rep("Muscle", 2), rep("Endocrine", 2), 
    rep("Specialized Epithelial", 2), rep("Glandular Epithelial", 5), 
    rep("Germ", 2), rep("Squamous Epithelial", 2), 
    rep("Trophoblast", 2), rep("Mesenchymal", 2)
  )
)
row_annotation$Tissue <- factor(row_annotation$Tissue, levels = c("Brain Cell Types", "Non-Brain Cell Types"))
row_annotation$Cat_Detailed <- factor(row_annotation$Cat_Detailed, levels = unique(row_annotation$Cat_Detailed))

## COLUMN ANNOTATIONS
column_annotation <- data.frame(
  Trait = c(
    rep("Blood Assays", 5), 
    rep("Reproductive and\n Activity Factors", 2), 
    rep("Physical\n Measure", 4), 
    rep("Cognitive\n Metrics", 4), 
    rep("Mental\n Health", 5)
  )
)
column_annotation$Trait <- factor(column_annotation$Trait, levels = unique(column_annotation$Trait))

column_ha <- HeatmapAnnotation(
  Trait = column_annotation$Trait,
  col = list(Trait = c(
    "Blood Assays" = "#e78ac3", 
    "Reproductive and\n Activity Factors" = "#8da0cb", 
    "Physical\n Measure" = "#a6d854", 
    "Cognitive\n Metrics" = "#66c2a5", 
    "Mental\n Health" = "#fc8d62"
  )),
  annotation_name = NULL, 
  show_legend = FALSE, 
  show_annotation_name = FALSE
)

# ==============================================================================
# 7. HEATMAP CONSTRUCTION
# ==============================================================================
num_detailed_splits <- length(unique(row_annotation$Cat_Detailed)) # 15
num_brain_categories <- 2

# Define the gap sizes (adjust these unit values as needed)
small_gap <- unit(0.2, "cm") # Separation between detailed categories
large_gap <- unit(1.2, "cm") # Larger separation between Brain and Non-Brain

# Create the vector of gaps (14 gaps for 15 groups)
row_gaps_list_raw <- list(
  # Gaps within Brain
  rep(small_gap, num_brain_categories - 1), 
  # Large gap between Brain/Non-Brain
  large_gap,                                
  # Gaps within Non-Brain
  rep(small_gap, num_detailed_splits - num_brain_categories - 1) 
)

# Flatten the list and combine all unit objects
row_gap_units <- do.call(unit.c, row_gaps_list_raw)

ht1 <- Heatmap(
  as.matrix(Matrix_HPACellType_Del_EffectSize_ordered), 
  name = "Effect size",
  na_col = "white",  
  
  cluster_rows = FALSE,  
  cluster_columns = FALSE,  
  show_row_dend = FALSE,  
  show_column_dend = FALSE,  
  
  top_annotation = column_ha,
  
  row_split = row_annotation$Cat_Detailed, 
  column_split = column_annotation$Trait,  
  
  row_title_rot = 0, 
  row_title_gp = gpar(fontsize = 8, fontface = "bold", fontfamily = "sans"),
  
  row_names_gp = gpar(col = "black", fontsize = 12, fontfamily = "sans", fontface = "bold"),
  column_title_gp = gpar(col = c('#e78ac3', '#8da0cb', '#a6d854', '#66c2a5', '#fc8d62'), fontsize = 14, fontface = "bold"),
  column_names_gp = gpar(col = "black", fontsize = 12, rot = 45, fontfamily = "sans", fontface = "bold"),
  column_names_rot = 45,
  
  cell_fun = function(j, i, x, y, width, height, fill) {
    if (!is.na(Matrix_HPACellType_Del_fdrpval_ordered[i, j])) {
      grid.text(
        Matrix_HPACellType_Del_fdrpval_ordered[i, j], 
        x = x, 
        y = y - convertHeight(grobHeight(textGrob('*')), "mm"), 
        gp = gpar(fontsize = 36, col = 'grey40', fontfamily = "sans", just = "center", fontface = "bold", hjust=30)
      )
    }
  },
  
  col = color_function, 
  row_gap = row_gap_units,   
  column_gap = unit(0.5, "cm"), 
  show_heatmap_legend = FALSE, 
  rect_gp = gpar(col = "lightgrey", lwd = 0.5) 
)


draw(ht1, padding = unit(c(0.5, 2, 0.5, 0.5), "cm"))

```




#### Panel B: Proportion of Associations/Enrichmenets  ####
##### Description: Differences in aggregate association/enrichment levels across variant types (Deletions, Duplications, SNPs) for all cell type gene sets. ####

##### Panel B Data ######

```{r}
# Load Del/Dup/SNP comparison stats 
load('df_compare_del_dup_snp.RData')

# Render a clean markdown table preview for GitHub
kable(head(df_compare_del_dup_snp), format = "markdown")
```


##### Panel B Figure ######

```{r , fig.width=8, fig.height=5}

# ------------------------------------------------------------------------------
# 1. DEFINE SAMPLE SIZE (N)
# ------------------------------------------------------------------------------
# Based on the HPA matrix dimensions (43 traits * 81 cell types)
total_pvalues_HPA <- 43 * 81

# ------------------------------------------------------------------------------
# 2. EXTRACT DATA & COMPUTE STANDARD ERROR
# ------------------------------------------------------------------------------
# IMPORTANT: Adjust 'HPA' if your dataframe uses a different exact string
df_hpa_props <- bind_rows(
  df_compare_del_dup_snp %>% 
    filter(Geneset_Type == "HPA") %>% 
    select(Type = Type1, prop = estimate_prop_Type1),
  df_compare_del_dup_snp %>% 
    filter(Geneset_Type == "HPA") %>% 
    select(Type = Type2, prop = estimate_prop_Type2)
) %>% 
  distinct() %>% 
  mutate(
    se = sqrt((prop * (1 - prop)) / total_pvalues_HPA),
    Type = factor(Type, levels = c("Del", "Dup", "SNP"))
  ) %>%
  arrange(Type)

# ------------------------------------------------------------------------------
# 3. AESTHETICS & COLOR PALETTE
# ------------------------------------------------------------------------------
custom_colors <- c(
  "Del" = "#bd3122",   
  "Dup" = "royalblue", 
  "SNP" = "#2e6320"    
)

# ------------------------------------------------------------------------------
# 4. GENERATE VISUALIZATION
# ------------------------------------------------------------------------------
p_panel_b <- ggplot(df_hpa_props, aes(x = Type, y = prop, fill = Type)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(
    aes(ymin = prop - se, ymax = prop + se), 
    width = 0.15,      
    linewidth = 0.8,        
    color = "black"    
  ) +
  scale_fill_manual(values = custom_colors) +
  labs(
    y = "Proportion of\nAssociations / Enrichments", 
    x = NULL 
  ) + 
  scale_y_continuous(
    limits = c(0, 0.25), 
    breaks = seq(0, 0.25, by = 0.05), 
    expand = c(0, 0)
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position   = "none",                                  
    axis.line         = element_line(linewidth = 1, color = "black"), 
    axis.ticks        = element_line(linewidth = 1, color = "black"), 
    axis.ticks.length = unit(0.2, "cm"),                         
    axis.text.y       = element_text(size = 16, color = "black", face = "bold"),
    axis.text.x       = element_text(size = 16, color = "black", face = "bold"),
    axis.title.y      = element_text(size = 16, face = "bold", margin = margin(r = 10))
  )

print(p_panel_b)

```


##### Panel B Stats ######

```{r , fig.width=6, fig.height=4}

formatted_stats_table <- df_compare_del_dup_snp %>%
  filter(Geneset_Type == "HPA") %>%
  select(
    Geneset_Type,
    Type1,
    Type2,
    p_value,
    statistic,
    estimate_prop_Type1,
    estimate_prop_Type2,
    p_value_fdr
  ) %>%
  mutate(
    statistic = round(statistic, 6),
    estimate_prop_Type1 = round(estimate_prop_Type1, 7),
    estimate_prop_Type2 = round(estimate_prop_Type2, 7),
    p_value = format(p_value, scientific = FALSE, digits = 7),
    p_value_fdr = format(p_value_fdr, scientific = FALSE, digits = 7)
  )

kable(
  formatted_stats_table, 
  align = "c",
  caption = "Pairwise Proportion Comparisons (HPA CellType81)"
)
```


#### Panel C: Comparison of Brain vs. Non-Brain Associations #### 
##### Description  This section generates a faceted analysis of the Cell type geneset, comparing the proportion of significant phenotypic associations between Brain and Non-Brain Cell-types The analysis is separated by structural variant type (Deletions, Duplications) and SNPs. ##### 


##### Panel C Data ######

```{r}
# Load Del/Dup/SNP comparison stats 
load('df_compare_brain_nonbrain.RData')

# Render a clean markdown table preview for GitHub
kable(head(df_compare_brain_nonbrain), format = "markdown")
```

##### Panel C Figure ######

```{r , fig.width=10, fig.height=5}
# ------------------------------------------------------------------------------
# 1. DEFINE SAMPLE SIZES (N)
# ------------------------------------------------------------------------------
# HPA Total: 81 Cell Types * 43 Traits
# Brain Tissues: 7 Cell Types * 43 Traits = 301
# Non-Brain Tissues: 74 Cell Types * 43 Traits = 3182
n_brain <- 7 * 43
n_nonbrain <- 74 * 43

# ------------------------------------------------------------------------------
# 2. EXTRACT DATA & COMPUTE STANDARD ERROR FOR PLOTTING
# ------------------------------------------------------------------------------
# IMPORTANT: Adjust 'HPA' if your dataframe uses a different exact string
df_panel_c_props <- bind_rows(
  df_compare_brain_nonbrain %>% 
    filter(Geneset_Type == "HPA") %>% 
    select(Type, Tissue_Class = geneset_cat1, prop = estimate_prop_cat1),
  df_compare_brain_nonbrain %>% 
    filter(Geneset_Type == "HPA") %>% 
    select(Type, Tissue_Class = geneset_cat2, prop = estimate_prop_cat2)
) %>% 
  distinct() %>% 
  mutate(
    N = case_when(
      Tissue_Class == "Brain"    ~ n_brain,
      Tissue_Class == "NonBrain" ~ n_nonbrain
    ),
    se = sqrt((prop * (1 - prop)) / N),
    Type = factor(Type, levels = c("Del", "Dup", "SNP")),
    Tissue_Class = factor(Tissue_Class, levels = c("Brain", "NonBrain"))
  ) %>%
  arrange(Type, Tissue_Class)

# ------------------------------------------------------------------------------
# 3. AESTHETICS & COLOR PALETTE
# ------------------------------------------------------------------------------
custom_colors_c <- c(
  "Brain"    = "#cf6323", 
  "NonBrain" = "#659e38"  
)

# ------------------------------------------------------------------------------
# 4. GENERATE VISUALIZATION
# ------------------------------------------------------------------------------
p_panel_c <- ggplot(df_panel_c_props, aes(x = Tissue_Class, y = prop, fill = Tissue_Class)) +
  geom_bar(stat = "identity", width = 0.4) +
  geom_errorbar(
    aes(ymin = prop - se, ymax = prop + se), 
    width = 0.1,      
    linewidth = 0.7,        
    color = "black"    
  ) +
  scale_fill_manual(values = custom_colors_c) +
  facet_wrap(~ Type) +
  labs(
    y = "Proportion of\nAssociations / Enrichments", 
    x = "Genesets" 
  ) + 
  scale_y_continuous(
    limits = c(0, 0.45),      
    breaks = seq(0, 0.4, by = 0.1), 
    expand = c(0, 0)
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position   = "none",                                  
    axis.line         = element_line(linewidth = 1, color = "black"), 
    axis.ticks        = element_line(linewidth = 1, color = "black"), 
    axis.ticks.length = unit(0.2, "cm"),                         
    axis.text.y       = element_text(size = 14, color = "black", face = "bold"),
    axis.text.x       = element_text(size = 14, color = "black", face = "bold"),
    axis.title.y      = element_text(size = 16, face = "bold", margin = margin(r = 10)),
    axis.title.x      = element_text(size = 16, face = "bold", margin = margin(t = 10)),
    strip.text        = element_text(size = 18, face = "bold"),  
    strip.background  = element_blank()                          
  )

print(p_panel_c)
```

##### Panel C Stats ######
```{r , fig.width=6, fig.height=4}
formatted_stats_c <- df_compare_brain_nonbrain %>%
  filter(Geneset_Type == "HPA") %>%
  select(
    Type,                 
    Geneset_Type,         
    Tissue_1 = geneset_cat1, 
    Tissue_2 = geneset_cat2, 
    p_value,
    statistic,
    estimate_prop_cat1,
    estimate_prop_cat2,
    p_value_fdr
  ) %>%
  mutate(
    statistic = round(statistic, 6),
    estimate_prop_cat1 = round(estimate_prop_cat1, 7),
    estimate_prop_cat2 = round(estimate_prop_cat2, 7),
    p_value = format(p_value, scientific = FALSE, digits = 7),
    p_value_fdr = format(p_value_fdr, scientific = FALSE, digits = 7)
  )

kable(
  formatted_stats_c, 
  align = "c",
  caption = "Brain vs Non-Brain Proportion Comparisons (HPA CellType81)"
)
```

#### Panel D: FunBurd on All of Us Cohort #### 
##### Description:  BMI Concordance Scatter Plots (UKBB vs AoU)##### 


##### Panel D Data (AoU) ######

```{r}
# Load AoU output
load('df_funburd_aou.RData')

kable(head(df_funburd_aou), format = "markdown")
```

##### Panel D Data (UKBB) ######
```{r}
# Load AoU output
load('df_funburd_ukbb.RData')

kable(head(df_funburd_ukbb), format = "markdown")
```

##### Panel D Figure ######
```{r , fig.width=10, fig.height=5}
target_traits <- c("BMI", "StandingHeight", "HeartRate", "Platelet")

df_panel_d_data <- inner_join(
  df_funburd_ukbb, 
  df_funburd_aou, 
  by = c("Trait", "Geneset", "Geneset_Type", "CNV_Type")
) %>%
  rename(Effectsize_UKBB = Effectsize) %>%
  mutate(
    UKBBLogp = -log10(p_value),
    AoULogp  = -log10(p_value_AoU),
    CNV_Type = factor(CNV_Type, levels = c("DEL", "DUP", "SNP"))
  ) %>%
  filter(Trait %in% target_traits) %>%
  mutate(Trait = factor(Trait, levels = target_traits))

# IMPORTANT: Adjust 'CellType_HPA81' if your dataframe uses a different exact string
df_bmi_concordance <- df_panel_d_data %>% 
  filter(Trait == "BMI", Geneset_Type == "SC_HPA81")

plot_concordance <- function(data, variant_type, title_text) {
  
  sub_df <- data %>% 
    filter(CNV_Type == variant_type) %>%
    filter(!is.na(Effectsize_UKBB) & !is.na(Effectsize_AoU))
  
  cor_res <- cor.test(sub_df$Effectsize_UKBB, sub_df$Effectsize_AoU)
  r_val <- cor_res$estimate
  p_val <- cor_res$p.value
  
  r_label <- paste0("r = ", round(r_val, 2), if(p_val < 0.05) "*" else "")
  
  ggplot(sub_df, aes(x = Effectsize_UKBB, y = Effectsize_AoU)) +
    geom_hline(yintercept = 0, color = "grey70", linewidth = 0.8) +
    geom_vline(xintercept = 0, color = "grey70", linewidth = 0.8) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60", linewidth = 1) +
    geom_point(color = "grey70", size = 2.5, alpha = 0.6) +
    geom_smooth(method = "lm", formula = y ~ x,color = "#E69F00", fill = "#E69F00", alpha = 0.2, linewidth = 1.5) +
    annotate("text", x = -Inf, y = Inf, label = r_label, 
             hjust = -0.1, vjust = 1.5, size = 6, fontface = "bold") +
    labs(title = title_text, x = "UKBB", y = "AoU") +
    theme_bw(base_size = 14) +
    theme(
      aspect.ratio = 1,
      panel.grid = element_blank(),
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.text = element_text(color = "black", size = 12),
      axis.title = element_text(face = "bold", size = 14)
    )
}

p_del <- plot_concordance(df_bmi_concordance, "DEL", "BMI\nDel EffectSize")
p_dup <- plot_concordance(df_bmi_concordance, "DUP", "BMI\nDup EffectSize")

p_panel_d_final <- p_del + p_dup
print(p_panel_d_final)
```


##### Panel D Stats ######

```{r , fig.width=10, fig.height=5}
formatted_stats_d <- df_panel_d_data %>%
  filter(Trait == "BMI", Geneset_Type == "SC_HPA81") %>%
  filter(!is.na(Effectsize_UKBB) & !is.na(Effectsize_AoU)) %>%
  group_by(Geneset_Type, Trait, CNV_Type) %>%
  filter(n() >= 3) %>%
  summarize(
    `Pearson r` = round(cor.test(Effectsize_UKBB, Effectsize_AoU)$estimate, 3),
    `Corr P-Value` = format(cor.test(Effectsize_UKBB, Effectsize_AoU)$p.value, scientific = TRUE, digits = 4),
    `Slope (β)` = round(coef(lm(Effectsize_AoU ~ Effectsize_UKBB))[2], 3),
    `Intercept` = round(coef(lm(Effectsize_AoU ~ Effectsize_UKBB))[1], 4),
    `R-Squared` = round(summary(lm(Effectsize_AoU ~ Effectsize_UKBB))$r.squared, 3),
    `N Variants` = n(),
    .groups = "drop"
  ) %>%
  arrange(Geneset_Type, Trait, CNV_Type)

kable(
  formatted_stats_d, 
  align = "c",
  caption = "Linear Regression & Correlation Stats: UKBB vs AoU Effect Sizes"
)
```


#### Panel E: Power vs Reproducibility (UKBB & AoU) #### 
##### Description:  Evaluates the relationship between discovery power (number of significant genesets in UKBB) and cross-cohort reproducibility (Pearson correlation of effect sizes between UKBB and AoU) ##### 

##### Panel E Figure ######

```{r , fig.width=7, fig.height=7}
df_concordance <- df_panel_d_data %>%
  # Using grepl to catch "HPA" or "CellType_HPA81"
  filter(grepl("HPA", Geneset_Type, ignore.case = TRUE)) %>%
  filter(CNV_Type %in% c("DEL", "DUP")) %>%
  group_by(Trait, CNV_Type) %>%
  summarise(
    Concordance = cor(Effectsize_UKBB, Effectsize_AoU, use = "complete.obs"), 
    .groups = "drop"
  )

df_sig_counts <- df_funburd_ukbb %>%
  filter(grepl("HPA", Geneset_Type, ignore.case = TRUE)) %>%
  filter(CNV_Type %in% c("DEL", "DUP")) %>%
  filter(FDR_p_value < 0.05) %>%
  group_by(Trait, CNV_Type) %>%
  summarise(N_Sig = n(), .groups = "drop")

df_plot <- df_concordance %>%
  left_join(df_sig_counts, by = c("Trait", "CNV_Type")) %>%
  mutate(
    N_Sig = ifelse(is.na(N_Sig), 0, N_Sig),
    Label = paste0(str_to_title(CNV_Type), "-", Trait)
  )

cor_test_res <- cor.test(df_plot$Concordance, df_plot$N_Sig)
global_r <- round(cor_test_res$estimate, 2)
global_p <- cor_test_res$p.value

global_label <- paste0("r = ", global_r, ifelse(global_p < 0.05, "*", ""))

p_summary <- ggplot(df_plot, aes(x = N_Sig, y = Concordance)) +
  geom_smooth(method = "lm",formula = y ~ x, color = "darkgrey", linetype = "solid", se = FALSE, linewidth = 1.2) +
  geom_point(aes(color = CNV_Type), size = 5) +
  geom_text_repel(aes(label = Label), 
                  size = 4.5, 
                  fontface = "bold", 
                  box.padding = 0.5,
                  point.padding = 0.3,
                  segment.color = "grey50",
                  color = "black") +
  annotate("text", x = Inf, y = Inf, label = global_label, 
           hjust = 1.1, vjust = 1.5, size = 6, fontface = "bold") +
  scale_color_manual(values = c("DEL" = "red2", "DUP" = "royalblue")) +
  labs(
    x = "N Sig. Genesets (UKBB)",
    y = "Effect-size Concordance (UKBB vs AoU)"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  coord_cartesian(ylim = c(-0.2, 1.05), clip = "off") +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 0.8),
    axis.ticks = element_line(colour = "black", linewidth = 0.8),
    axis.text = element_text(size = 12, face = "bold", color = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.position = "none",
    plot.margin = margin(20, 20, 20, 20)
  )

print(p_summary)
```

##### Panel E Stats ######

```{r , fig.width=10, fig.height=5}
formatted_stats_e <- df_plot %>%
  filter(!is.na(N_Sig) & !is.na(Concordance)) %>%
  mutate(Scope = "4 Mutual Traits") %>%
  group_by(Scope) %>%
  filter(n() >= 3) %>%
  summarize(
    `Pearson r` = round(cor.test(N_Sig, Concordance)$estimate, 3),
    `Corr P-Value` = format(cor.test(N_Sig, Concordance)$p.value, scientific = TRUE, digits = 4),
    `Slope (β)` = round(coef(lm(Concordance ~ N_Sig))[2], 4),
    `Intercept` = round(coef(lm(Concordance ~ N_Sig))[1], 4),
    `R-Squared` = round(summary(lm(Concordance ~ N_Sig))$r.squared, 3),
    `N Data Points` = n(),
    .groups = "drop"
  )

kable(
  formatted_stats_e, 
  align = "c",
  caption = "Linear Regression & Correlation Stats: Discovery Power vs. Reproducibility"
)

```